Estimating duration using survival analysis

### 2023-10-04 cis-perfect-testing.tex
Updates from Overleaf

The estimation of incidence from prevalence is sensitive to the tail of the duration distribution.
In \cref{E-ATACCC} I estimated the duration distribution from the ATACCC dataset.
However, this study suffered from a small sample size and only 20 days of follow-up.
Therefore, model assumptions determined the size of the tail estimated.
In this chapter, I develop a method to estimate the distribution using CIS data. 
CIS sampling is less frequent, but with a large sample size and long follow-up.

@@

The CIS tests each individual infrequently, either weekly or monthly (see\todo{refer to section with CIS intro}).
The infrequent testing complicates analysis of the data in two ways.
First, infection episodes can go undetected.
Second, the infection episode start and end times are known imprecisely.
This section explains these issues.
The rest of the chapter develops a method to model them.

The first complication in the analysis is that infection episodes can be undetected.
This means there are no positive tests associated with them.
Undetected episodes occur when an episode starts and ends between tests (see \cref{perf-test:fig:truncation}).
For undetected episodes, there is no information on them episode; even the number that occurred is unknown.
Undetected episodes are important because they are shorter than detected episodes.
To see why this is the case, consider the undetected infection in the bottom of \cref{perf-test:fig:truncation}.
The infection episode depicted begins at time 33 and has a 10-day duration.
If the duration was 24 days or more, the individual would still be detectable at their next test and detected then.
CIS has many undetected infections. This is because frequency of testing (up to 28 days, see\todo{ref section on the survey design}) is longer than most infections (median around 15 days).
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{cis-perfect-testing/truncation}}
  \caption[Undetected episodes in CIS data]{An unknown number of infection episodes in CIS are undetected. Consider the two individuals shown here: we collect identical data for them (a series of negative tests) yet the top individual was never infected and the bottom individual was. See the main text for why this is an issue. \label{perf-test:fig:truncation}}
\end{figure}

@@

I base my analysis on the 4,800 episodes that were first detected between 10th Oct 2020 and 6th Dec 2020 inclusive with negatives bounding the start and end time of the episode.
That is, the individual tested negative before and after the episode.
Without a negative before the episode there is no lower bound on the episode's start time.
It could be arbitrarily long and there is little information on its length.
A negative test follows each episode unless the individual was lost to follow-up before their episode ended.
This is rare.
Episodes removed in this step are incorporated in the adjustment for undetected episodes.

>>>>

The estimation of incidence from prevalence is sensitive to the tail of the duration distribution.
In \cref{E-ATACCC} I estimated the duration distribution from the ATACCC dataset.
However, this study suffered from a small sample size and only 20 days of follow-up.
Therefore, model assumptions determined the size of the estimated tail estimated.
In this chapter, I develop a method to estimate the duration distribution using CIS data. CIS sampling is less frequent, but with a large sample size and long follow-up.

@@

In the CIS each individualis tested  infrequently, either weekly or monthly (see\todo{refer to section with CIS intro}).
The infrequent testing complicates analysis of the data in two ways.
First, infection episodes can go undetected.
Second, the infection episode start and end times are known imprecisely.
This section explains these issues. The rest of the chapter develops a method to tackle  them.

The first complication in the analysis is that infection episodes can be undetected.
This means there are no positive tests associated with them.
Undetected episodes occur when an episode starts and ends between tests (see \cref{perf-test:fig:truncation}).
For undetected episodes, there is no information on them episode; even the number that occurred is unknown.
Undetected episodes are important because they are shorter than detected episodes.
To see why this is the case, consider the undetected infection in the bottom of \cref{perf-test:fig:truncation}.
The infection episode depicted begins at time 33 and has a 10-day duration.
If the duration had been 24 days or more, the individual would still be detectable at their next test and detected then.
CIS has many undetected infections. This is because frequency of testing (up to 28 days, see\todo{ref section on the survey design}) is longer than most infections (median around 15 days).
\begin{figure}
  \makebox[\textwidth][c]{\includegraphics[width=1.2\textwidth]{cis-perfect-testing/truncation}}
  \caption[Undetected episodes in CIS data]{An unknown number of infection episodes in CIS are undetected. Consider the two individuals shown here: we collect identical data for them (a series of negative tests) yet the top individual was never infected and the bottom individual was. See the main text for why this is an issue. \label{perf-test:fig:truncation}}
\end{figure}

@@

I base my analysis on the 4,800\todo{check number} episodes that were first detected between 10th Oct 2020 and 6th Dec 2020 inclusive with negatives bounding the start and end time of the episode.
That is, the individual tested negative before and after the episode.
Without a negative before the episode there is no lower bound on the episode's start time.
It could be arbitrarily long and there is little information on its length.
A negative test follows each episode unless the individual was lost to follow-up before their episode ended.
This is rare.
Episodes removed in this step are incorporated in the adjustment for undetected episodes.
